---
title: "p8105_hw2_jl6049"
author: "LYU JING"
date: "10/2/2021"
output: github_document
---

## Problem 1

Read and clean the Mr. Trash Wheel sheet:

```{r}
library(tidyverse)
library(readxl)
Trash_data = 
  read_excel("../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>%
  janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls = round(sports_balls))


```


specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year.
```{r}

Pre_2019_data = 
  read_excel("../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", skip = 1)%>%
  janitor::clean_names()%>%
  drop_na(total,month)%>%
  mutate(year = 2019)


Pre_2018_data = 
  read_excel("../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1)%>%
  janitor::clean_names()%>%
  drop_na(total,month)%>%
  mutate(year = 2018)
```

 Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}



Pre_combine = 
  bind_rows(Pre_2018_data, Pre_2019_data)%>%
  mutate(month = month.name[month])%>%
  select(year, month, total)


summary(Trash_data)
skimr::skim(Trash_data)%>%
  select(numeric.mean)

```

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?

For the first sheet "Mr. Trash Wheel", we could got `r nrow(Trash_data)` observations of dumpster


